Value Probability
0 p
1 2p
2 3p
3 4p
4 5p
5 6p
Table 4.4: The Distribution of Y

Question 4.1. Table 4.6 presents the probabilities of the random variable Y .
These probabilities are a function of the number p, the probability of the value
“0”. Answer the following questions:
1. What is the value of p = 1/21
2. P(Y < 3) = 6/21
3. P(Y = odd) = 12/21
4. P(1 ≤ Y < 4) = 9/21
5. P(|Y − 3| < 1.5) = 12/21
6. E(Y ) = 3.333
7. Var(Y ) = 2.222
8. What is the standard deviation of Y = sqrt(2.22)


Question 4.2. One invests $2 to participate in a game of chance. In this game
a coin is tossed three times. If all tosses end up “Head” then the player wins
$10. Otherwise, the player losses the investment.
1. What is the probability of winning the game?

(0.5 * 0.5 * 0.5) = 1/8 

2. What is the probability of loosing the game?

1 - 1/8 = 7/8

3. What is the expected gain for the player that plays this game? (Notice that the expectation can obtain a negative value.)

E[gain] = 8 * 1/8 + (-2) * (7/8)


Question 6.1. Consider the problem of establishing regulations concerning the
maximum number of people who can occupy a lift. In particular, we would like
to assess the probability of exceeding maximal weight when 8 people are allowed
to use the lift simultaneously and compare that to the probability of allowing 9
people into the lift.
Assume that the total weight of 8 people chosen at random follows a normal
distribution with a mean of 560kg and a standard deviation of 57kg. Assume
that the total weight of 9 people chosen at random follows a normal distribution
with a mean of 630kg and a standard deviation of 61kg.
1. What is the probability that the total weight of 8 people exceeds 650kg?

```{r}

val <- pnorm(650,560,57)
1 - val

# 0.05717406

```

2. What is the probability that the total weight of 9 people exceeds 650kg?

```{r}

val <- pnorm(650,630,61)
1 - val

# 0.3715054

```

3. What is the central region that contains 80% of distribution of the total
weight of 8 people?

```{r}

val1 <- qnorm(0.1, 560, 57)
val1
val2 <- qnorm(0.9, 560, 57)
val2

# [486.9516, 633.0484]

```

4. What is the central region that contains 80% of distribution of the total
weight of 9 people?

```{r}

val3 <- qnorm(0.1, 630, 61)
val3
val4 <- qnorm(0.9, 630, 61)
val4

# [551.8254, 708.1746]

```

Question 7.1. The file “pop2.csv” contains information associated to the
blood pressure of an imaginary population of size 100,000. The file can be found
on the internet (http://pluto.huji.ac.il/~msby/StatThink/Datasets/pop2.
csv). The variables in this file are:
id: A numerical variable. A 7 digits number that serves as a unique identifier
of the subject.
sex: A factor variable. The sex of each subject. The values are either “MALE”
or “FEMALE”.
age: A numerical variable. The age of each subject.
bmi: A numerical variable. The body mass index of each subject.
systolic: A numerical variable. The systolic blood pressure of each subject.
diastolic: A numerical variable. The diastolic blood pressure of each subject.
group: A factor variable. The blood pressure category of each subject. The
values are “NORMAL” both the systolic blood pressure is within its normal
range (between 90 and 139) and the diastolic blood pressure is within its
normal range (between 60 and 89). The value is “HIGH” if either measurements of blood pressure are above their normal upper limits and it is
“LOW” if either measurements are below their normal lower limits.
Our goal in this question is to investigate the sampling distribution of the sample
average of the variable “bmi”. We assume a sample of size n = 150.
1. Compute the population average of the variable “bmi”.

```{r}

pop2 <- read.csv("C:\\Users\\ds3\\Downloads\\pop2.csv")
head(pop2)

```

2. Compute the population standard deviation of the variable “bmi”
3. Compute the expectation of the sampling distribution for the sample average of the variable.
4. Compute the standard deviation of the sampling distribution for the sample average of the variable.
5. Identify, using simulations, the central region that contains 80% of the
sampling distribution of the sample average.
6. Identify, using the Central Limit Theorem, an approximation of the central region that contains 80% of the sampling distribution of the sample
average.


