# Question 5.20 

Body measurements, Part III. Exercise 5.13 introduces data on shoulder girth and
height of a group of individuals. The mean shoulder girth is 108.20 cm with a standard deviation
of 10.37 cm. The mean height is 171.14 cm with a standard deviation of 9.41 cm. The correlation
between height and shoulder girth is 0.67.

(a) Write the equation of the regression line for predicting height.

slope = r * (sy/sx)
slope = 0.67 * (9.41/10.37)

b = y - mx
b = 171.14 - slope(108.20)

```{r}

slope <- 0.67 * 9.41 / 10.37
slope
# 0.6079749

intercept <- 171.14 - slope * 108.20
intercept
# 105.3571

```

y = 0.6079749x + 105.3571

(b) Interpret the slope and the intercept in this context.

The slope means that for every cm increase in shoulder girth, height will increase 
~0.6079 cm. 

(c) Calculate R^2 of the regression line for predicting height from shoulder girth, and interpret it
in the context of the application.

```{r}

r_squared <- slope ^ 2
r_squared

```

The R^2 is 0.3696335 which indicates that ~37% of the variability in height can be explained
by shoulder girth.

(d) A randomly selected student from your class has a shoulder girth of 100 cm. Predict the height
of this student using the model.

```{r}

predicted_girth <- 0.6079749 * 100 + 105.3571
predicted_girth
#  166.1546 cm

```

(e) The student from part (d) is 160 cm tall. Calculate the residual, and explain what this residual
means.

```{r}

residual <- intercept - predicted_girth
residual
# -60.79748 cm

```
The residual means that the actual height of the person is ~60.8 cm shorter than the
predicted height based on the shoulder girth of 100 cm.


(f) A one year old has a shoulder girth of 56 cm. Would it be appropriate to use this linear model
to predict the height of this child?

No, it would not make sense becuase the sample is taken within a certain
range (likely for adults) and therefore a child with a small girth would not
be appropriate to fit into this model. 



# Question 5.29 

Body measurements, Part IV. The scatterplot and least squares summary below show
the relationship between weight measured in kilograms and height measured in centimeters of 507
physically active individuals.

Estimate Std. Error t value Pr(>|t|)
(Intercept) -105.0113 7.5394 -13.93 0.0000
height 1.0176 0.0440 23.13 0.0000

(a) Describe the relationship between height and weight.

There is a strong, positive linear correlation between height and weight.

(b) Write the equation of the regression line. Interpret the slope and intercept in context.

weight = 1.0176x - 105.0113

The intercept is the predicted weight when height is 0. The slope means that for
every cm in height, weight will increase by 1.0176 kg. 


(c) Do the data provide strong evidence that an increase in height is associated with an increase
in weight? State the null and alternative hypotheses, report the p-value, and state your
conclusion.

Ho = There is no relationship between height and weight
Ha = There is a relationship between height and weight

The p-value is ~0 meaning that we can reject the null hypothesis and conclude
that there is a relationship between height and weight.

(d) The correlation coefficient for height and weight is 0.72. Calculate R^2
and interpret it in context.

```{r}

r_sqrd <- 0.72 ^ 2
r_sqrd
# 0.5184

```

About 51.84% of the variability in weight can be explained by 
height using the regression model.