# Question 12.1. 

Consider a medical condition that does not have a standard
treatment. The recommended design of a clinical trial for a new treatment
to such condition involves using a placebo treatment as a control. A placebo
treatment is a treatment that externally looks identical to the actual treatment
but, in reality, it does not have the active ingredients. The reason for using
placebo for control is the “placebo effect”. Patients tent to react to the fact that
they are being treated regardless of the actual beneficial effect of the treatment.
As an example, consider the trial for testing magnets as a treatment for pain
that was described in Question 9.1. The patients that where randomly assigned
to the control (the last 21 observations in the file “magnets.csv”) were treated
with devises that looked like magnets but actually were not. The goal in this
exercise is to test for the presence of a placebo effect in the case study “Magnets
and Pain Relief” of Question 9.1 using the data in the file “magnets.csv”.

1. Let X be the measurement of change, the difference between the score of
pain before the treatment and the score after the treatment, for patients
that were treated with the inactive placebo. Express, in terms of the
expected value of X, the null hypothesis and the alternative hypothesis
for a statistical test to determine the presence of a placebo effect. The null
hypothesis should reflect the situation that the placebo effect is absent.

Ho: E(x) = 0 (mean change in pain score for plecebo group is 0)
Ha: E(x) =/ 0 (mean chnage in pain score for plecevo group is not 0)


2. Identify the observations that can be used in order to test the hypotheses.

The last 21 oberservations is the group that was tested with the plecebo.

3. Carry out the test and report your conclusion. (Use a significance level of
5%.)

```{r}

magnets <- read.csv("C:\\Users\\ds3\\Downloads\\magnets.csv")
head(magnets)

placebo_data <- magnets |> slice(30:50) |> pull(change)
t_test_result <- t.test(placebo_data, mu = 0)
t_test_result

#         One Sample t-test

# data:  placebo_data
# t = 3.1804, df = 20, p-value = 0.004702
# alternative hypothesis: true mean is not equal to 0
# 95 percent confidence interval:
#  0.3768845 1.8135916
# sample estimates:
# mean of x
#  1.095238

t_test_result$p.value < 0.05
# TRUE

```

Conclusion: Reject the null hypothesis. There is evidence of a placebo effect.


# Question 13.1. 

In this exercise we would like to analyze the results of the
trial that involves magnets as a treatment for pain. The trial is described in
Question 9.1. The results of the trial are provided in the file “magnets.csv”.
Patients in this trail where randomly assigned to a treatment or to a control.
The responses relevant for this analysis are either the variable “change”, which
measures the difference in the score of pain reported by the patients before and
after the treatment, or the variable “score1”, which measures the score of pain
before a device is applied. The explanatory variable is the factor “active”.
This factor has two levels, level “1” to indicate the application of an active
magnet and level “2” to indicate the application of an inactive placebo.
In the following questions you are required to carry out tests of hypotheses.
All tests should conducted at the 5% significance level:

1. Is there a significance difference between the treatment and the control
groups in the expectation of the reported score of pain before the application of the device?

Ho: μ1 = μ2 (There is no difference between the mean pain score of the treatment and the control group)
Ha: μ1 =/ μ2 (There is a difference between the mean pain score of the treatment and the control group)


```{r}

magnets <- read.csv("C:\\Users\\ds3\\Downloads\\magnets.csv")
head(magnets)

t_test_score1 <- t.test(score1 ~ active, data = magnets)
t_test_score1

# Welch Two Sample t-test

# data:  score1 by active
# t = 0.41483, df = 38.273, p-value = 0.6806
# alternative hypothesis: true difference in means between group "1" and group "2" is not equal to 0
# 95 percent confidence interval:
#  -0.3757896  0.5695498
# sample estimates:
# mean in group "1" mean in group "2" 
#           9.62069           9.52381

t_test_score1$p.value
# 0.6805825

```

The mean test statistic is not stitistically significant and therefore we cannot reject the null 
hypothesis.

2. Is there a significance difference between the treatment and the control
groups in the variance of the reported score of pain before the application
of the device?

Ho: μ1 = μ2 (There is no difference between the variance pain score of the treatment and the control group)
Ha: μ1 =/ μ2 (There is a difference between the variance pain score of the treatment and the control group)

```{r}

var_test_score1 <- var.test(score1 ~ active, data = magnets)
var_test_score1

#         F test to compare two variances

# data:  score1 by active
# F = 0.69504, num df = 28, denom df = 20, p-value = 0.3687
# alternative hypothesis: true ratio of variances is not equal to 1
# 95 percent confidence interval:
#  0.2938038 1.5516218
# sample estimates:
# ratio of variances
#          0.6950431

var_test_score1$p.value

```

The variance test statistic is not statistically significant and therefore we cannot reject the null
hypothesis.

3. Is there a significance difference between the treatment and the control
groups in the expectation of the change in score that resulted from the
application of the device?

Ho: μ1 = μ2 (There is no mean difference between the change of the pain score)
Ha: μ1 =/ μ2 (There is a mean difference between the change of the pain score)


```{r}

t_test_change <- t.test(change ~ active, data = magnets)
t_test_change
t_test_change$p.value
# 3.860128e-07

```

There change test statistic is statistically significent and therefore we can reject the null
hypothesis

4. Is there a significance difference between the treatment and the control
groups in the variance of the change in score that resulted from the application of the device?

Ho: μ1 = μ2 (There is no variance difference between the change of the pain score)
Ha: μ1 =/ μ2 (There is a variance difference between the change of the pain score)

```{r}

var_test_change <- var.test(change ~ active, data = magnets)
var_test_change

#         F test to compare two variances

# data:  change by active
# F = 4.2062, num df = 28, denom df = 20, p-value = 0.001535
# alternative hypothesis: true ratio of variances is not equal to 1
# 95 percent confidence interval:
#  1.778003 9.389902
# sample estimates:
# ratio of variances
#           4.206171

```

The variance test statistic is stitistically significant and therefore we can reject the null
hypothesis.
