# Question 9.1. 

The data for the 50 patients is stored in file “magnets.csv”.
The file can be found on the internet at http://pluto.huji.ac.il/~msby/
StatThink/Datasets/magnets.csv. Download this file to your computer and
store it in the working directory of R. Read the content of the file into an R data
frame. Produce a summary of the content of the data frame and answer the
following questions:

```{r}
library(tidyverse)
library(dplyr)

magnets <- read.csv("C:\\Users\\ds3\\Downloads\\magnets.csv")
summary(magnets)
head(magnets)

```


1. What is the sample average of the change in score between the patient’s
rating before the application of the device and the rating after the application?

Mean = 3.5


2. Is the variable “active” a factor or a numeric variable?

`active` is a factor


3. Compute the average value of the variable “change” for the patients that
received and active magnet and average value for those that received an
inactive placebo. (Hint: Notice that the first 29 patients received an active
magnet and the last 21 patients received an inactive placebo. The subsequence of the first 29 values of the given variables can be obtained via
the expression “change[1:29]” and the last 21 vales are obtained via the
expression “change[30:50]”.)

```{r}

magnets <- read.csv("C:\\Users\\ds3\\Downloads\\magnets.csv")

magnets |> slice(1:29) |> pull(change) |> mean()
magnets |> slice(30:50) |> pull(change) |> mean()

```


4. Compute the sample standard deviation of the variable “change” for the
patients that received and active magnet and the sample standard deviation for those that received an inactive placebo.

```{r}

magnets |> slice(1:29) |> pull(change) |> sd()
magnets |> slice(30:50) |> pull(change) |> sd()

```

5. Produce a boxplot of the variable “change” for the patients that received
and active magnet and for patients that received an inactive placebo.
What is the number of outliers in each subsequence?

```{r}

magnets |>
    slice(1:29) |>
    pull(change) |>
    boxplot()

magnets |>
    slice(30:50) |>
    pull(change) |>
    boxplot()

```

Active magnet had no outliers, while inactive placebo group has 3 outliers



# Question 10.1. 

In Subsection 10.3.2 we compare the average against the midrange as estimators of the expectation of the measurement. The goal of this
exercise is to repeat the analysis, but this time compare the average to the
median as estimators of the expectation in symmetric distributions.


1. Simulate the sampling distribution of average and the median of a sample
of size n = 100 from the Normal(3, 2) distribution. Compute the expectation and the variance of the sample average and of the sample median.
Which of the two estimators has a smaller mean square error?

```{r}

sample_means <- replicate(1e5, mean(rnorm(100, 3, sqrt(2))))
sample_medians <- replicate(1e5, median(rnorm(100, 3, sqrt(2))))

mean_mean <- mean(sample_means)
mean_mean
# 2.999894
var_mean <- var(sample_means)
var_mean
# 0.0198959
mse_mean <- mean((sample_means - 3)^2)
mse_mean
#0.01989572

mean_median <- mean(sample_medians)
mean_median
# 2.999761
var_median <- var(sample_medians)
var_median
# 0.03117613
mse_median <- mean((sample_medians - 3)^2)
mse_median
# 0.03117588

```


2. Simulate the sampling distribution of average and the median of a sample
of size n = 100 from the Uniform(0.5, 5.5) distribution. Compute the
expectation and the variance of the sample average and of the sample
median. Which of the two estimators has a smaller mean square error?


```{r}

mu <- (a + b) / 2

sample_means <- replicate(1e5, mean(runif(100, 0.5, 5.5)))
sample_medians <- replicate(1e5, median(runif(100, 0.5, 5.5)))

# Compute statistics
mean_mean <- mean(sample_means)
var_mean <- var(sample_means)
mse_mean <- mean((sample_means - mu)^2)

mean_median <- mean(sample_medians)
var_median <- var(sample_medians)
mse_median <- mean((sample_medians - mu)^2)

```


# Question 10.2. 

The goal in this exercise is to assess estimation of a proportion
in a population on the basis of the proportion in the sample.
The file “pop2.csv” was introduced in Exercise 7.1 of Chapter 7. This file
contains information associated to the blood pressure of an imaginary population of size 100,000. The file can be found on the internet (http://pluto.
huji.ac.il/~msby/StatThink/Datasets/pop2.csv). One of the variables in
the file is a factor by the name “group” that identifies levels of blood pressure.
The levels of this variable are “HIGH”, “LOW”, and “NORMAL”.
The file “ex2.csv” contains a sample of size n = 150 taken from the given
population. This file can also be found on the internet (http://pluto.huji.
ac.il/~msby/StatThink/Datasets/ex2.csv). It contains the same variables
as in the file “pop2.csv”. The file “ex2.csv” corresponds in this exercise to
the observed sample and the file “pop2.csv” corresponds to the unobserved
population.
Download both files to your computer and answer the following questions:


1. Compute the proportion in the sample of those with a high level of blood
pressure16

```{r}

pop2 <- read.csv("C:\\Users\\ds3\\Downloads\\pop2.csv")
head(pop2)

ex2 <- read.csv("C:\\Users\\ds3\\Downloads\\ex2.csv")
head(ex2)


high_bp <- filter(ex2, group == "HIGH") |> nrow()
total_sample <- filter(ex2, group %in% c("HIGH", "LOW", "NORMAL")) |> nrow()
high_bp / total_sample
# 0.2466667

```

2. Compute the proportion in the population of those with a high level of
blood pressure.

```{r}

high_bp_pop <- filter(pop2, group == "HIGH") |> nrow()
total_pop <- filter(pop2, group %in% c("HIGH", "LOW", "NORMAL")) |> nrow()
high_bp_pop / total_pop
# 0.28126

```

3. Simulate the sampling distribution of the sample proportion and compute
its expectation.

```{r}


sample_proportions <- replicate(
  1e5,
  {
    sample_data <- sample_n(pop2, 150)
    group_column <- pull(sample_data, group)
    mean(group_column == "HIGH")
  }
)

mean(sample_proportions)
# 0.2812202

```

4. Compute the variance of the sample proportion.

```{r}

var(sample_proportions)
# 0.001335008

```

5. It is proposed in Section 10.5 that the variance of the sample proportion
is Var(Pˆ) = p(1 − p)/n, where p is the probability of the event (having a
high blood pressure in our case) and n is the sample size (n = 150 in our
case). Examine this proposal in the current setting.

```{r}

p <- 0.2812202
n <- 150

theoretical_var <- p * (1 - p) / n
theoretical_var
# 0.001347569

```


# Question 2.2. 

Heart transplants, Part II. Exercise 1.50 introduces the Stanford Heart Transplant Study.
Of the 34 patients in the control group, 4 were alive at the end of the study. Of the 69 patients in
the treatment group, 24 were alive. The contingency table below summarizes these results.
Group
Control Treatment Total
Alive 4 24 28 Outcome Dead 30 45 75
Total 34 69 103

(a) What proportion of patients in the treatment group and what proportion of patients in the
control group died?

45/69
30/34

(b) One approach for investigating whether or not the treatment is effective is to use a randomization technique.
i. What are the claims being tested? Use the same null and alternative hypothesis notation
used in the section.

Null Hypothesis (H0): The treatment has no effect --> pc = pt
Alternative Hypothesis (Ha): The treatment is effective --> pc > pt

ii. The paragraph below describes the set up for such approach, if we were to do it without using statistical software. Fill in the blanks with a number or phrase, whichever is
appropriate.
We write alive __28__on cards representing patients who were alive at
the end of the study, and dead on __75___ cards representing patients
who were not. Then, we shuffle these cards and split them into two groups:
one group of size __64__representing treatment, and another group of
size __34__representing control. We calculate the difference between
the proportion of dead cards in the treatment and control groups (treatment -
control) and record this value. We repeat this many times to build a distribution
centered at pt-pc=0. Lastly, we calculate the fraction of simulations where
the simulated differences in proportions are pt-pc>0. If this fraction is low,
we conclude that it is unlikely to have observed such an outcome by chance and
that the null hypothesis should be rejected in favor of the alternative.

iii. What do the simulation results shown below suggest about the effectiveness of the transplant program?

41/99 show that the treatment was effective. Fail to reject the null hypothesis.


# Question 2.6 

Is yawning contagious, Part I. An experiment conducted by the MythBusters, a science
entertainment TV program on the Discovery Channel, tested if a person can be subconsciously
influenced into yawning if another person near them yawns. 50 people were randomly assigned to
two groups: 34 to a group where a person near them yawned (treatment) and 16 to a group where
there wasn’t a person yawning near them (control). 

The following table shows the results of this experiment.

Group
Treatment Control Total
Result Yawn 10 4 14
Not Yawn 24 12 36
Total 34 16 50

A simulation was conducted to understand the distribution of the test statistic under the assumption of independence: having someone yawn near another person has no influence on if the other
person will yawn. In order to conduct the simulation, a researcher wrote yawn on 14 index cards
and not yawn on 36 index cards to indicate whether or not a person yawned. Then he shuffled
the cards and dealt them into two groups of size 34 and 16 for treatment and control, respectively.
He counted how many participants in each simulated group yawned in an apparent response to a
nearby yawning person, and calculated the difference between the simulated proportions of yawning as ˆptrtmt,sim − pˆctrl,sim. This simulation was repeated 10,000 times using software to obtain
10,000 differences that are due to chance alone. The histogram shows the distribution of the
simulated differences.

(a) What are the hypotheses?

Null Hypothesis (H0): Yawning is not contagious --> pt = pc
Alternative Hypothesis (Ha): The treatment is effective --> pc < pt

(b) Calculate the observed difference between the yawning rates under the two scenarios.

10/34 - 4/16


(c) Estimate the p-value using the figure above and determine the conclusion of the hypothesis
test.

```{r}

result <- 0.001 + 0.02 + 0.07 + 0.17 + 0.245 + 0.24 + 0.155 + 0.065 + 0.025 + 0.01 + 0.0005
result

result <- 0.24 + 0.155 + 0.065 + 0.025 + 0.01 + 0.0005
result

```

We do not have enough evidence to prove that yawning is contageous


# Question 3.1 

Vegetarian college students. Suppose that 8% of college students are vegetarians. Determine if the following statements are true or false, and explain your reasoning.
(a) The distribution of the sample proportions of vegetarians in random samples of size 60 is
approximately normal since n ≥ 30.

False - it does not meet the second condition of np>=10 or n(1-p)>=10

(b) The distribution of the sample proportions of vegetarian college students in random samples
of size 50 is right skewed.

True - it is right skewed 

(c) A random sample of 125 college students where 12% are vegetarians would be considered
unusual.

(d) A random sample of 250 college students where 12% are vegetarians would be considered
unusual.
(e) The standard error would be reduced by one-half if we increased the sample size from 125
to 250.