
# Question 5.29

See if you can reproduce the table in ISRS 5.29 using the original dataset in body.dat.txt.

```{r}

body_data <- read.table("C:\\Users\\ds3\\coursework\\week3\\body.dat.txt", header = FALSE)
head(body_data)

colnames(body_data)[c(23, 24)] <- c("weight", "height")

model <- lm(weight ~ height, data = body_data)
summary(model)

# Coefficients:
#               Estimate Std. Error t value Pr(>|t|)    
# (Intercept) -105.01125    7.53941  -13.93   <2e-16 ***
# height         1.01762    0.04399   23.14   <2e-16 ***

```

# Question 6.1 

Baby weights, Part I. The Child Health and Development Studies investigate a range of
topics. One study considered all pregnancies between 1960 and 1967 among women in the Kaiser
Foundation Health Plan in the San Francisco East Bay area. Here, we study the relationship
between smoking and weight of the baby. The variable smoke is coded 1 if the mother is a
smoker, and 0 if not. The summary table below shows the results of a linear regression model for
predicting the average birth weight of babies, measured in ounces, based on the smoking status of
the mother.

Estimate Std. Error t value Pr(>|t|)
(Intercept) 123.05 0.65 189.60 0.0000
smoke -8.94 1.03 -8.65 0.0000

The variability within the smokers and non-smokers are about equal and the distributions are
symmetric. With these conditions satisfied, it is reasonable to apply the model. (Note that we
don’t need to check linearity since the predictor has only two levels.)

```{r}

birth_weight_data <- read.table("C:\\Users\\ds3\\coursework\\week3\\babyweights.txt")
head(birth_weight_data)

model <- lm(bwt ~ smoke, data = birth_weight_data)
summary(model)

# Coefficients:
#             Estimate Std. Error t value Pr(>|t|)
# (Intercept)  123.047      0.649 189.597   <2e-16 ***
# smoke         -8.938      1.033  -8.653   <2e-16 ***

```

(a) Write the equation of the regression line.

y = -8.94x + 123.05

(b) Interpret the slope in this context, and calculate the predicted birth weight of babies born to
smoker and non-smoker mothers.

The slope indicates that babies born to smokers weigh 8.94 oz less than those 
born to non-smokers.

Smoker:
```{r}
predicted_weight_sm <- -8.94 * 1 + 123.05
predicted_weight_sm
# 114.11
```

Non-smoker
```{r}
predicted_weight_ns <- -8.94 * 0 + 123.05
predicted_weight_ns
# 123.05
```


(c) Is there a statistically significant relationship between the average birth weight and smoking?

The p-value is ~0 meaning that the relationship between smoking 
and birth weight is statistically significant.


# Question 6.2 

Baby weights, Part II. Exercise 6.1 introduces a data set on birth weight of babies.
Another variable we consider is parity, which is 0 if the child is the first born, and 1 otherwise.
The summary table below shows the results of a linear regression model for predicting the average
birth weight of babies, measured in ounces, from parity.

Estimate Std. Error t value Pr(>|t|)
(Intercept) 120.07 0.60 199.94 0.0000
parity -1.93 1.19 -1.62 0.1052

```{r}

birth_weight_data <- read.table("C:\\Users\\ds3\\coursework\\week3\\babyweights.txt")
head(birth_weight_data)

model <- lm(bwt ~ parity, data = birth_weight_data)
summary(model)

# Coefficients:
#             Estimate Std. Error t value Pr(>|t|)    
# (Intercept) 120.0684     0.6005 199.942   <2e-16 ***
# parity       -1.9287     1.1895  -1.621    0.105   

```

(a) Write the equation of the regression line.

y = -1.9287x + 120.0684

(b) Interpret the slope in this context, and calculate the predicted birth weight of first borns and
others.

The slope indicates that babies that are not firstborn weigh 1.9287 oz less than
firstborn babies on average. 

Firstborn:
```{r}
predicted_weight_f <- -1.9287 * 0 + 120.0684
predicted_weight_f
# 120.0684
```

Non-firstborn
```{r}
predicted_weight_nf <- -1.9287 * 1 + 120.0684
predicted_weight_nf
# 118.1397
```

(c) Is there a statistically significant relationship between the average birth weight and parity?

The p-value is greater than 0.1 indicating that there is not enough evidence 
to conclude that there is a statistically significant relationship between
parity and birth weight. 


# Question 6.3 

Baby weights, Part III. We considered the variables smoke and parity, one at a time, in
modeling birth weights of babies in Exercises 6.1 and 6.2. A more realistic approach to modeling
infant weights is to consider all possibly related variables at once. Other variables of interest
include length of pregnancy in days (gestation), mother’s age in years (age), mother’s height in
inches (height), and mother’s pregnancy weight in pounds (weight). Below are three observations
from this data set.

bwt gestation parity age height weight smoke
1 120 284 0 27 62 100 0
2 113 282 0 33 64 135 0
.. .. .. .. .. .. .. ..
1236 117 297 0 38 65 129 0

The summary table below shows the results of a regression model for predicting the average birth
weight of babies based on all of the variables included in the data set.

Estimate Std. Error t value Pr(>|t|)
(Intercept) -80.41 14.35 -5.60 0.0000
gestation 0.44 0.03 15.26 0.0000
parity -3.33 1.13 -2.95 0.0033
age -0.01 0.09 -0.10 0.9170
height 1.15 0.21 5.63 0.0000
weight 0.05 0.03 1.99 0.0471
smoke -8.40 0.95 -8.81 0.0000

```{r}

model_full <- lm(bwt ~ gestation + parity + age + height + weight + smoke, data = birth_weight_data)
summary(model_full)

# Coefficients:
#              Estimate Std. Error t value Pr(>|t|)    
# (Intercept) -80.41085   14.34657  -5.605 2.60e-08 ***
# gestation     0.44398    0.02910  15.258  < 2e-16 ***
# parity       -3.32720    1.12895  -2.947  0.00327 ** 
# age          -0.00895    0.08582  -0.104  0.91696
# height        1.15402    0.20502   5.629 2.27e-08 ***
# weight        0.05017    0.02524   1.987  0.04711 *
# smoke        -8.40073    0.95382  -8.807  < 2e-16 **

```

(a) Write the equation of the regression line that includes all of the variables.

bwt = −80.41 + (0.44 * gestation) − (3.33 * parity) − (0.01 * age) + (1.15 * height) + (0.05 * weight) − (8.40 * smoke)

(b) Interpret the slopes of gestation and age in this context.

gestation: For each additional day of pregnancy, the birth weight increases by 0.44 oz.
age: For each additional year of the mother's age, the birth weight decreases by
     0.01 oz.

(c) The coefficient for parity is different than in the linear model shown in Exercise 6.2. Why
might there be a difference?

Before, parity was the only predictor. Now there are other variables in play and therefore
the effect of the parity variable is adjusted. 

(d) Calculate the residual for the first observation in the data set.

```{r}

predicted_first <- predict(model_full, newdata = birth_weight_data[1, ])
actual_first <- birth_weight_data$bwt[1]

residual_first <- actual_first - predicted_first
residual_first
# -2.003102

```


(e) The variance of the residuals is 249.28, and the variance of the birth weights of all babies
in the data set is 332.57. Calculate the R^2 and the adjusted R^2
. Note that there are 1,236
observations in the data set.

```{r}

residuals_var <- var(resid(model_full))
total_var <- var(birth_weight_data$bwt)

r_squared <- 1 - (residuals_var / total_var)
r_squared
# 0.2504296

n <- nrow(birth_weight_data)
k <- length(model_full$coefficients) - 1
adj_r_squared <- 1 - ((1 - r_squared) * (n - 1) / (n - k - 1))
adj_r_squared
# 0.2467702

```